\thetitle

### 2021-03-03 writing/mutate-benchmarks.tex
bad latex
mutate-*: rewording

   \itme  fail and contain type-level mistakes. 
\end{itemize}

@@

Hence, even if they do not come with type-level errors, they offer a
diverse lignuistic substrate suitable for the careful introduction of
type-level errors via mutations. Further filtering the benchmark suite to
keep the benchmarks with the most sizeable and complex dependency graphs
yields ten suitable programs while preserving the suite's linguistic
diversity. Figure~\ref{table:benchmark-descriptions} displays the selected
benchmark programs and describes their basic characteristic.

>>>>

   \item  fail and contain type-level mistakes. 
\end{itemize}

@@

Hence, even if they do not come with type-level errors, they offer a
diverse linguistic substrate suitable for the careful introduction of
type-level errors via mutations. Further filtering the benchmark suite to
keep the benchmarks with the most sizeable and complex dependency graphs
yields ten suitable programs while preserving the suite's linguistic
diversity. Figure~\ref{table:benchmark-descriptions} displays the selected
benchmark programs and describes their basic characteristics.

### 2021-03-03 writing/mutate-interesting.tex
mutate-*: rewording

A type-level mutation is {\em interesting\/} (1) if, upon removal of the type
annotations, running the program raises a run-time error, (2) whose stack trace
contains source locations from at least three modules, and (3) if this happens
in all of the to-be-evaluated semantics. An interesting mutation represents a
{\em debugging scenario\/} that supplies the rational programmer with
information.

@@

\item By definition, an impedance mismatch is a clash between the type ascription of
one module's imports and another module's exports. Hence, type checking fails for an
interesting mutant module, and removing the types from this mutant module
creates an impedance mismatch between it and the modules that import it,
directly or indirectly. This impedance mismatch is of interest if, and only if,
running the program leads to an exception---otherwise it goes unobserved and
cannot be debugged.

\item If the evaluation od a mutant module immediately raises an exception because
of the changes, there is no work for the rational programmer. Indeed, if the
stack trace contains source pointers to two modules, the scenario is still
uninteresting. Every ordinary benchmark program comes with a {\tt main} module
that acts as a driver, whose source is guaranteed to be included in the stack
trace.  Hence, the definition of interesting mutation insists on the presence of
three different modules in the stack trace. This guarantees that the debugging
scenario demands a sufficiently sophisticated effort, due to the interaction
between the buggy module with its context.  In these cases, the rational
programmer must examine at least two modules to locate the source of the faulty
interaction.

\item The goal of a comparative evaluation is to give the rational programmer a
chance to debug the same scenario using different pieces of information.  In the
case of gradual typing semantics, a meta-theorem due to Greenman and
Felleisen~\cite{gf-icfp-2018} says that if a program raises an exception under
Erasure, it also errors under all other semantics.  Hence, a comparison of blame
information insists that a debugging scenario {\em raises a run-time exception
under Erasure\/}.

@@

@@

The definition of interesting debugging scenarios creates a powerful filter. All
together, the listed mutators produce 16,800 mutants with at least one
interesting debugging scenario across all benchmarks; see
figure~\ref{fig:mutant-breakdown} for an overview. Broken down by benchmark, the
mutators produce at least 40 interesting scenarios for every benchmark, and
these scenarios originate from at least four different mutators per benchmark.
Thus, the mutators result in a sizable and diverse population of scenarios for
every benchmark.  Furthermore, every mutator contributes scenarios to at least
one benchmark.  Some mutators apply only to a few benchmarks, because they
target rather specific features; for instance, the class-focused mutators are
mainly effective in a program that makes extensive use of object-oriented
features.

>>>>

A type-level mutation is {\em interesting\/} (1) if the type checker rejects the fully typed version of the mutant,
(2) running the mutant with all type annotations removed raises a run-time error, and
(3) that error's stack trace contains source locations from at least three modules.

@@

\item By definition, an impedance mismatch is a clash between the type ascription of one module's imports and another module's exports. Hence, type checking should fail for an interesting mutant.

\item The goal of a comparative evaluation is to give the rational programmer a
chance to debug the same scenario using different pieces of information.  In the
case of gradual typing semantics, a meta-theorem due to Greenman and
Felleisen~\cite{gf-icfp-2018} says that if a program raises an exception under
Erasure, it also errors under all other semantics.  Hence, a comparison of blame
information insists that an interesting mutant {\em raises a run-time exception
under Erasure\/}.

@@

\item If the evaluation of a mutant module immediately raises an exception because
of the changes, there is no work for the rational programmer. Indeed, if the
stack trace contains source pointers to two modules, the scenario is still
uninteresting. Every ordinary benchmark program comes with a {\tt main} module
that acts as a driver, whose source is guaranteed to be included in the stack
trace.  Hence, the definition of interesting mutation insists on the presence of
three different modules in the stack trace. This guarantees that the debugging
scenario demands a sufficiently sophisticated effort, due to the interaction
between the buggy module with its context.  In these cases, the rational
programmer must contend with at least two modules involved in a faulty
interaction.

@@

The definition of interesting mutants creates a powerful filter. All
together, the listed mutators produce 16,800 interesting mutants across all benchmarks; see
figure~\ref{fig:mutant-breakdown} for an overview. Broken down by benchmark, the
mutators produce at least 40 interesting mutants for every benchmark, and
these mutants originate from at least four different mutators per benchmark.
Thus, the mutators result in a sizable and diverse population of scenarios for
every benchmark.  Furthermore, every mutator contributes interesting mutants in at least
one benchmark.  Some mutators apply only to a few benchmarks, because they
target rather specific features; for instance, the class-focused mutators are
mainly effective in a program that makes extensive use of object-oriented
features.

The definition of interesting mutants is more than a filter for
uninteresting mutants. It is also the guide for evaluating and refining
mutators until they collectively produce a sizable and diverse population of
interesting mutants. For example, one early mutator idea casts the
tests of conditional expressions to the {\tt Any} type to
suppress occurrence typing. However, an examination of the mutants it
produces through the lens of the definition of interesting mutants
reveals that none is interesting because they never raise run-time errors.
Indeed, the interesting standard guided countless iterations of adding,
removing, and refining the mutators in figure~\ref{table:mutation-ops}
to maximize interesting mutants.

### 2021-03-03 writing/mutate-mutators.tex
small tweaks to emphasize interesting mutations
mutate-*: rewording

For the evaluation of a blame strategy, mutators must produce type-level
mistakes that the run-time checks of gradual typing systems or the safety
checks of the underlying language can detect. Once detected, the rational
programmer should be able to locate the mistake by gradually adding types to
blamed modules.

Figure~\ref{table:mutation-ops} describes 16 mutators that satisfy these
constraints. As the explanation indicates, some specialize or
generalize~\citet{lksfd-popl-2020}'s mutators, which in turn are borrowed from
the vast literature on mutation testing~\cite{jia2011analysis}.\footnote{Many
of~\citet{lksfd-popl-2020} mutators do not satisfy the constraints and had to
be eliminated. For example, the relational mutator does not lead to type-level
errors. Similarly, the arithmetic mutator does not reliably lead to type
errors; it had to be adapted to make only operator swaps that can affect the
type of the result, e.g., changing `*` to `/`.}  Only two are directly
inherited; many mutators are brand new. For the latter the authors relied on
their decades-long experience of making type-level mistakes in Typed Racket,
some of which take a non-trivial effort to debug.

Most of the mutators are self-explanatory.  The first six apply to all
gradually typed languages; the next six to those that include classes and
objects. The last four target distinguishing features of Typed Racket,
specifically is sophisticated type system. In particular, {\tt arithmetic}
replaces a {\tt +} with a {\tt -} in an attempt to change the type of the
arithmetic expression; {\tt +}'s result is a {\tt Positive-Integer} when all
arguments are positive integers, while {\tt -} yields {\tt
Integer}~\cite{stathff-padl-12}. Similarly, the other three also aim to usurp
Typed Racket's occurrence type system.

>>>>

For the evaluation of a blame strategy, mutators must produce type-level
mistakes that the run-time checks of gradual typing systems or the safety
checks of the underlying language can detect. Once detected, the rational
programmer should be able to locate the mistake by gradually adding types
to blamed modules. In other words, the suitability of the mutators hinges
on their ability to generate interesting debugging scenarios (see
sec.~\ref{sub:mutate-interesting}).

Figure~\ref{table:mutation-ops} describes 16 mutators that satisfy these
constraints. As the explanation indicates, some specialize or
generalize~\citet{lksfd-popl-2020}'s mutators, which in turn are borrowed from
the vast literature on mutation testing~\cite{jia2011analysis}.\footnote{Many
of~\citet{lksfd-popl-2020} mutators do not satisfy the constraints and had to
be eliminated or significantly adapted. For example, the relational mutator does not lead to type-level
errors. Similarly, the arithmetic mutator does not reliably lead to type
errors; it had to be adapted to make only operator swaps that can affect the
type of the result, e.g., changing `*` to `/`.}  Only two are directly
inherited; many mutators are brand new. For the latter the authors relied on
their decades-long experience of making type-level mistakes in Typed Racket,
some of which take non-trivial effort to debug.

Most of the mutators are self-explanatory.  The first six apply to all
gradually typed languages; the next six to those that include classes and
objects. The last four target distinguishing features of Typed Racket,
specifically its sophisticated type system. For example, one mutation produced by {\tt arithmetic}
replaces a {\tt +} with a {\tt -} in an attempt to change the type of the
arithmetic expression; {\tt +}'s result is a {\tt Positive-Integer} when all
arguments are positive integers, while {\tt -} yields {\tt
Integer}~\cite{stathff-padl-12}. Similarly, the other three aim to confound
Typed Racket's occurrence type system.

### 2021-03-03 writing/mutate.tex
re-established the old organization
editiroal voice thru 6
mutate-*: rewording

Putting a rational programmer to work means generating many mutants and turning
those into debugging scenarios. Since existing mutators do not generate useful
mutants, the first step is to develop new mutators (see sec.~\ref{sub:mutate-mutators})
and to validate their suitability for this application (see sec.~\ref{sub:mutate-interesting}).

@@

\input{mutate-mutators-figure.tex}
\sub{mutators}   {How to Mutate Software} 
\sub{interesting}{Are These Mutators Interesting}

>>>>

Putting the rational programmer to work means generating many mutants and turning
those into debugging scenarios. The process must start with a suitable
collection of representative programs (see sec.~\ref{sub:mutate-benchmarks}).
Since existing mutators do not generate useful mutants, the first step is to
develop new mutators (see sec.~\ref{sub:mutate-mutators}) and to validate their
suitability for this application (see sec.~\ref{sub:mutate-interesting}).

@@

\input{mutate-benchmarks-figure.tex}
\sub{benchmarks}{The Experimental Benchmarks}
\input{mutate-mutators-figure.tex}
\sub{mutators}   {How to Mutate Software} 
\sub{interesting}{Are These Mutators Interesting}
